Inter-task effects introduce bias in crowdsourcing

### 2014-08-03 docs/publications/2014.05_PNAS/manuscript.tex
fix selling

In the present work, we investigate inter-task priming using image-labeling 
tasks on the AMT platform.  
(***why we are using image labelling as prototypical task***) 
Each worker is required to label ten images 
featuring 
food and culture.  We regard these images as being made up of an 
\textit{initial} and \textit{final set}, but, crucially, there is no
distinction made between these sets, nor any interruption in the flow of tasks 
from the view of the worker.
Each worker is randomly assigned one of three initial sets, but the final set
of images is always the same.  We analyze the effect that the assignment of 
initial set has on workers' attribution of labels to the final set.  We 
measure effects using our novel approach, as well as looking for changes in 
the content of labels and label specificity based on an ontology we 
developped for that purpose.

@@

Surprisingly, we find that inter-task effects on label orientation and 
specificity are much stronger than those brought about by framing tasks with a 
funder name.  Our results show that initial tasks can introduce significant 
bias in subsequent worker labeling.  Interestingly, we find that inter-task
effects can be used to induce greater specificity.  Our results suggest that
workers attribute more specific labels when labelling a series of images that
that are more similar to one another.  This suggests that careful 
consideration should be given to the bundling of tasks when designing a study 
using a microtask platform.

Our analysis also highlights important advantages of our novel definition of 
priming.  Our approach relies on no prior knowledge about the nature
of the task task whatsoever.  In contrast, we also sought to detect priming
based on its direct effects on label content and specificity, implicitly
drawing on prior knowledge about the the images and label semantics.
Remarkably, while being more general, our novel approach was also more 
sensitive.  This suggests that our measurement approach could be applied in 
cases where one does not \emph{a priori} know whether any priming has occured, 
nor what the effects might be.

@@

In the language of computer science, $\mathcal{A}$ is a classifier performing
a binary classification.  The measure $\theta$ can range between 0 and 1, and 
indicates how much better the performance of the best possible classifier is 
compared to a random guess. We intentionally leave the details of 
$\mathcal{A}$ unrestricted because $\theta$ should depend on the pontential 
for two populations to be distinguished, based on their inherent differences, 
rather than on the performance of a particular classifier.

In practice, some particular classifier $\mathcal{C}$ must be used and a 
statistical estimate of its performance $\hat{\theta}_\mathcal{C}$ obtained. 
$\hat{\theta}_\mathcal{C}$ acts as a lower bound for the 
theoretical value $\theta$ from the definition, because there might 
exist some other classifier $\mathcal{C}'$ with 
$\hat{\theta}_\mathcal{C'}>\hat{\theta}_\mathcal{C}$.
Here, we use a Naive Bayes classifier because it simple and performs well 
accross a broad range of classification tasks[***].

To interpret $\hat{\theta}_\mathcal{C}$, imagine that some choice is to be
made between two options on the basis of worker outputs (e.g. whether or not 
to flag an image as innappropriate). Any such binary choice has a formal 
structure of a trained classifier (i.e. a distinguisher which discards the 
first two inputs). This means that $\hat{\theta}_\mathcal{C}$ estimates a 
lower bound for the worst-case bias introduced in such a binary decision when 
one failed to control for priming effects.

@@

%\begin{addmargin}[11em]{11em}
%	\vspace{2mm}
%	$^a$\footnotesize{The full funder names used were ``The Glodal Foundation
%		for Cultural Recognition'' and ``The National Foundation for 
%		Nutritional Awareness''.}
%\end{addmargin}

@@

With some exceptions, the classifier performed significantly better 
than chance when distinguishing treatments that had different initial image 
sets.  This demonstrates that inter-task effects were strong for the labelling 
tasks.  Of the sixteen treatment pairs having different initial images, there 
were three cases where the classifier performance was on par with chance. 
Each of these cases involved distinguishing workers shown the ambiguous set 
from ones shown the the ingredients set.  The reason for the low 
distinguishability of workers from these treatments might be because
the ambiguous and ingredients sets are somewhat similar to one another.
We will come back to this point later when we discuss the degree of similarity 
between image sets.

@@

\begin{figure}
	\includegraphics[scale=0.94]{figs/f1-thetas_full_pairwise25_img1.pdf}
	\caption{ \footnotesize{
		Empirical difference in priming, $\hat{\theta}_\text{NB}$, between 
		various treatments.  The difference in priming between each pair of 
		treatments 
		was determined by cross-validation of a Naive Bayes classifier trained
		to classify workers based on the labels they attributed to the first
		image of the final image set. Each panel compares a basis treatment,
		indicated above the pannel, to various treatments indicated on the
		abscissa.  We can conclude, with 95\% confidence, that the classifier
		is performing better than chance when $\hat{\theta}_\text{NB}$ 
		exceeds 0.168, indicated by the dotted line.  See \textbf{Methods} for
		the calculation of this confidence threshold.
	}}
\end{figure}

@@

We can similarly look at how the inter-task effects on label composition
(i.e. fraction of culture- and food-oriented words) play out as workers 
progress through the final image set.  To do this, we define the 
\textit{excess cultural orientation} as the number of culture-orientod labels 
minus the number of food oriented ones.  Tracking the excess cultural 
orientation across the images is, however, not in itself meaningful,
because some images will inherently carry more cultural content than others. 
To account for this, we calculate the excess cultural content for both 
\textsc{AMBG} and $\textsc{CULT}_{img}$, and subtract the former from the 
latter, yeilding the \textit{relative} excess cultural content, 
$\Delta_{cult}$.  Formally,

@@

In the present study, we are more interested in the ultimate outputs of 
workers' rather than the psychological details which lead to them (though 
these are interesting in their own right).  This allows us to sidestep the 
complication in attributing similarity: we asses how similar two sets of 
images are, 
\textit{with respect to the labelling task}, by looking at the fraction of 
labels that they share.  Formally, we compute the similarity of two images
sets $X$ and $Y$ by computing the Jacquard index between the sets of labels 
attributed to them:

@@

The similarities of the various image sets are presented in \textbf{Table 2}. 
First, observe that the ambiguous and ingredients sets are much more similar 
to eachother than either is to the culture set.  This is consistent with our
finding that $\hat{\theta}_\text{NB}$ between \textsc{AMBG} and 
$\textsc{INGR}_{img}$ was not significant (see \textbf{Fig 1}).

@@

This suggests that labelling a series of images that are similar to 
one another might encourage workers to use more specific terminology.
Such a phenomenon would be consistent with the psychological mechanism known 
as \textit{negative priming}.  Negative priming occurs when a person becomes 
desensitized to non-salient stimuli to which she is repeatedly 
exposed [***].  Consider that workers who initially 
labeled the ambiguous image set had already seen five images showing 
prepared meals once they labelled the first test image.  At that point,
a worker might not regard the labels \texttt{food} or \texttt{meal} to be 
salient, since they do not distinguish the image from those she had seen 
already.

@@

In one sense, this is not surprising because the Naive 
Bayes classifier takes into account the frequency of occurence of all labels 
encountered during training.  But it is remarkable that the 
method which encorporates neither prior knowledge about the image contents nor 
label semantics, outperforms those that do.  This is a powerful feature 
because, given the outputs from two sets of workers, one can test whether they 
have been differently primed without knowing how that priming might manifest.

@@

\paragraph{Batching tasks for better HPU performance.}
Our proposed connection between similarity and specificity during 
image-labelling might be used to tune the specificity of labels.  For example,
If one seeks very nuanced labelling, our results suggest that the images 
should first be sorted into batches based on their similarity. This could be 
accomplished by beginning with a first, coarse characterization of unsorted 
images, followed by batching based on the labels obtained. Then, the batches
of similar images could be served for a second round of finer labelling. 
The sorting and re-labelling could then in principle be repeated.

In effect, this creates a hierarchical workflow, where the first, 
coarsely-characterizing HPUs provide a stream of labelled images that is 
split (batched) and fed to other HPUs for finer characterization.  This 
workflow exhibits an interesting difference between HPUs and CPUs.  In general,
whenever an aspect of a problem can be parallelized when employing CPUs, one
gains efficiency.  But here, because of HPU hysteresis, we see an example where
one might sacrifice accuracy through parallelization.

@@

\textbf{Statistical significance of $\theta_\text{NB}$.}
We test the significance two population's measured $\theta_\text{NB}$ against 
the null hypothesis that that the two populations are in fact identical:
\begin{align}
	\mathbf{H_0}: \theta_\text{NB} = 0 \\
	\mathbf{H_1}: \theta_\text{NB} > 0
\end{align}
If the populations are in fact identical, then the probability that the 
classifier correctly classifies a worker is $\frac{1}{2}$. Let $K$ be a random
variable which denotes the number of correct classifications when the 
classifier is chalenged to classify a total of $n$ workers during 
cross-validation.  Under $\mathbf{H_0}$, the probability distribution of $K$ 
is:
\begin{align}
	\text{Pr}\{K = k | n \} = { n \choose k } \left(\frac{1}{2}\right)^n.
\end{align}
We can reject $\mathbf{H_0}$ at significance $\alpha$ if $\theta_\text{NB}$ 
is greater than or equal to a critical value, $\theta_\text{NB}^*$:
\begin{align}
	\theta^*_\text{NB} = \frac{2k^*}{n} - 1
\end{align}
where $k^*$ is the smallest integer in $\{0,\dots,n\}$ that satisfies 
\begin{align}
	\sum_{k \in \{k^*, \dots, n\}} \text{Pr}\{K = k | n \} < \alpha.
\end{align}
For a 5-fold cross-validation with non-overlapping test sets containing 25 
workers ($n = 125$), and a significance of $\alpha = 0.05$, we
have $k^*=73 \implies \theta_\text{NB}^* = 0.168$.

@@

\begin{figure}
	\includegraphics[scale=1.00]{figs/taskImages/ambiguous.jpg}
	\caption{ Ambiguous image set. These images were presented to workers from 
		certain treatements (see \textbf{Table 1}) in the order shown.}
\end{figure}

\begin{figure}
	\includegraphics[scale=1.00]{figs/taskImages/cultural.jpg}
	\caption{Cultural image set. These images were presented to workers from 
		certain treatements (see \textbf{Table 1}) in the order shown.}
\end{figure}

\begin{figure}
	\includegraphics[scale=1.00]{figs/taskImages/ingredients.jpg}
	\caption{ Ingredients image set. These images were presented to workers 
		from certain treatements (see \textbf{Table 1}) in the order shown.}
\end{figure}
\section*{Acknowledgements}
\section*{References}
\begingroup
\renewcommand{\chapter}[2]{}
\bibliography{newbib.bib}
\endgroup
\bibliographystyle{plain}

>>>>

In the present work, we investigate inter-task priming using image-labeling 
tasks on the AMT platform.  
(***why we are using image labeling as prototypical task***) 
Each worker is required to label ten images 
featuring 
food and culture.  We regard these images as being made up of an 
\textit{initial} and \textit{final set}, but, crucially, there is no
distinction made between these sets, nor any interruption in the flow of tasks 
from the view of the worker.
Each worker is randomly assigned one of three initial sets, but the final set
of images is always the same.  We analyze the effect that the assignment of 
initial set has on workers' attribution of labels to the final set.  We 
measure effects using our novel approach, as well as looking for changes in 
the content of labels and label specificity based on an ontology we 
developed for that purpose.

@@

Surprisingly, we find that inter-task effects on label orientation and 
specificity are much stronger than those brought about by framing tasks with a 
funder name.  Our results show that initial tasks can introduce significant 
bias in subsequent worker labeling.  Interestingly, we find that inter-task
effects can be used to induce greater specificity.  Our results suggest that
workers attribute more specific labels when labeling a series of images that
that are more similar to one another.  This suggests that careful 
consideration should be given to the bundling of tasks when designing a study 
using a microtask platform.

Our analysis also highlights important advantages of our novel definition of 
priming.  Our approach relies on no prior knowledge about the nature
of the task task whatsoever.  In contrast, we also sought to detect priming
based on its direct effects on label content and specificity, implicitly
drawing on prior knowledge about the images and label semantics.
Remarkably, while being more general, our novel approach was also more 
sensitive.  This suggests that our measurement approach could be applied in 
cases where one does not \emph{a priori} know whether any priming has occurred, 
nor what the effects might be.

@@

In the language of computer science, $\mathcal{A}$ is a classifier performing
a binary classification.  The measure $\theta$ can range between 0 and 1, and 
indicates how much better the performance of the best possible classifier is 
compared to a random guess. We intentionally leave the details of 
$\mathcal{A}$ unrestricted because $\theta$ should depend on the potential 
for two populations to be distinguished, based on their inherent differences, 
rather than on the performance of a particular classifier.

In practice, some particular classifier $\mathcal{C}$ must be used and a 
statistical estimate of its performance $\hat{\theta}_\mathcal{C}$ obtained. 
$\hat{\theta}_\mathcal{C}$ acts as a lower bound for the 
theoretical value $\theta$ from the definition, because there might 
exist some other classifier $\mathcal{C}'$ with 
$\hat{\theta}_\mathcal{C'}>\hat{\theta}_\mathcal{C}$.
Here, we use a Naive Bayes classifier because it simple and performs well 
across a broad range of classification tasks[***].

To interpret $\hat{\theta}_\mathcal{C}$, imagine that some choice is to be
made between two options on the basis of worker outputs (e.g. whether or not 
to flag an image as inappropriate). Any such binary choice has a formal 
structure of a trained classifier (i.e. a distinguisher which discards the 
first two inputs). This means that $\hat{\theta}_\mathcal{C}$ estimates a 
lower bound for the worst-case bias introduced in such a binary decision when 
one failed to control for priming effects.

@@

@@

With some exceptions, the classifier performed significantly better 
than chance when distinguishing treatments that had different initial image 
sets.  This demonstrates that inter-task effects were strong for the labeling 
tasks.  Of the sixteen treatment pairs having different initial images, there 
were three cases where the classifier performance was on par with chance. 
Each of these cases involved distinguishing workers shown the ambiguous set 
from ones shown the ingredients set.  The reason for the low 
distinguishability of workers from these treatments might be because
the ambiguous and ingredients sets are somewhat similar to one another.
We will come back to this point later when we discuss the degree of similarity 
between image sets.

@@

\begin{figure}
	\includegraphics[scale=0.94]{figs/f1-thetas_full_pairwise25_img1.pdf}
	\caption{ \footnotesize{
		Empirical difference in priming, $\hat{\theta}_\text{NB}$, between 
		various treatments.  The difference in priming between each pair of 
		treatments 
		was determined by cross-validation of a Naive Bayes classifier trained
		to classify workers based on the labels they attributed to the first
		image of the final image set. Each panel compares a basis treatment,
		indicated above the panel, to various treatments indicated on the
		abscissa.  We can conclude, with 95\% confidence, that the classifier
		is performing better than chance when $\hat{\theta}_\text{NB}$ 
		exceeds 0.168, indicated by the dotted line.  See \textbf{Methods} for
		the calculation of this confidence threshold.
	}}
\end{figure}

@@

We can similarly look at how the inter-task effects on label composition
(i.e. fraction of culture- and food-oriented words) play out as workers 
progress through the final image set.  To do this, we define the 
\textit{excess cultural orientation} as the number of culture-oriented labels 
minus the number of food oriented ones.  Tracking the excess cultural 
orientation across the images is, however, not in itself meaningful,
because some images will inherently carry more cultural content than others. 
To account for this, we calculate the excess cultural content for both 
\textsc{AMBG} and $\textsc{CULT}_{img}$, and subtract the former from the 
latter, yielding the \textit{relative} excess cultural content, 
$\Delta_{cult}$.  Formally,

@@

In the present study, we are more interested in the ultimate outputs of 
workers' rather than the psychological details which lead to them (though 
these are interesting in their own right).  This allows us to sidestep the 
complication in attributing similarity: we asses how similar two sets of 
images are, 
\textit{with respect to the labeling task}, by looking at the fraction of 
labels that they share.  Formally, we compute the similarity of two images
sets $X$ and $Y$ by computing the Jacquard index between the sets of labels 
attributed to them:

@@

The similarities of the various image sets are presented in \textbf{Table 2}. 
First, observe that the ambiguous and ingredients sets are much more similar 
to each other than either is to the culture set.  This is consistent with our
finding that $\hat{\theta}_\text{NB}$ between \textsc{AMBG} and 
$\textsc{INGR}_{img}$ was not significant (see \textbf{Fig 1}).

@@

This suggests that labeling a series of images that are similar to 
one another might encourage workers to use more specific terminology.
Such a phenomenon would be consistent with the psychological mechanism known 
as \textit{negative priming}.  Negative priming occurs when a person becomes 
desensitized to non-salient stimuli to which she is repeatedly 
exposed [***].  Consider that workers who initially 
labeled the ambiguous image set had already seen five images showing 
prepared meals once they labeled the first test image.  At that point,
a worker might not regard the labels \texttt{food} or \texttt{meal} to be 
salient, since they do not distinguish the image from those she had seen 
already.

@@

In one sense, this is not surprising because the Naive 
Bayes classifier takes into account the frequency of occurrence of all labels 
encountered during training.  But it is remarkable that the 
method which incorporates neither prior knowledge about the image contents nor 
label semantics, outperforms those that do.  This is a powerful feature 
because, given the outputs from two sets of workers, one can test whether they 
have been differently primed without knowing how that priming might manifest.

@@

\paragraph{Batching tasks for better HPU performance.}
Our proposed connection between similarity and specificity during 
image-labeling might be used to tune the specificity of labels.  For example,
If one seeks very nuanced labeling, our results suggest that the images 
should first be sorted into batches based on their similarity. This could be 
accomplished by beginning with a first, coarse characterization of unsorted 
images, followed by batching based on the labels obtained. Then, the batches
of similar images could be served for a second round of finer labeling. 
The sorting and re-labeling could then in principle be repeated.

In effect, this creates a hierarchical workflow, where the first, 
coarsely-characterizing HPUs provide a stream of labeled images that is 
split (batched) and fed to other HPUs for finer characterization.  This 
workflow exhibits an interesting difference between HPUs and CPUs.  In general,
whenever an aspect of a problem can be parallelized when employing CPUs, one
gains efficiency.  But here, because of HPU hysteresis, we see an example where
one might sacrifice accuracy through parallelization.

@@

\textbf{Statistical significance of $\theta_\text{NB}$.}
We test the significance two population's measured $\theta_\text{NB}$ against 
the null hypothesis that that the two populations are in fact identical:
\begin{align}
	\mathbf{H_0}: \theta_\text{NB} = 0 \\
	\mathbf{H_1}: \theta_\text{NB} > 0
\end{align}
If the populations are in fact identical, then the probability that the 
classifier correctly classifies a worker is $\frac{1}{2}$. Let $K$ be a random
variable which denotes the number of correct classifications when the 
classifier is challenged to classify a total of $n$ workers during 
cross-validation.  Under $\mathbf{H_0}$, the probability distribution of $K$ 
is:
\begin{align}
	\text{Pr}\{K = k | n \} = { n \choose k } \left(\frac{1}{2}\right)^n.
\end{align}
We can reject $\mathbf{H_0}$ at significance $\alpha$ if $\theta_\text{NB}$ 
is greater than or equal to a critical value, $\theta_\text{NB}^*$:
\begin{align}
	\theta^*_\text{NB} = \frac{2k^*}{n} - 1
\end{align}
where $k^*$ is the smallest integer in $\{0,\dots,n\}$ that satisfies 
\begin{align}
	\sum_{k \in \{k^*, \dots, n\}} \text{Pr}\{K = k | n \} < \alpha.
\end{align}
For a 5-fold cross-validation with non-overlapping test sets containing 25 
workers ($n = 125$), and a significance of $\alpha = 0.05$, we
have $k^*=73 \implies \theta_\text{NB}^* = 0.168$.

@@

\begin{figure}
	\includegraphics[scale=1.00]{figs/taskImages/ambiguous.jpg}
	\caption{ Ambiguous image set. These images were presented to workers from 
		certain treatments (see \textbf{Table 1}) in the order shown.}
\end{figure}

\begin{figure}
	\includegraphics[scale=1.00]{figs/taskImages/cultural.jpg}
	\caption{Cultural image set. These images were presented to workers from 
		certain treatments (see \textbf{Table 1}) in the order shown.}
\end{figure}

\begin{figure}
	\includegraphics[scale=1.00]{figs/taskImages/ingredients.jpg}
	\caption{ Ingredients image set. These images were presented to workers 
		from certain treatments (see \textbf{Table 1}) in the order shown.}
\end{figure}
\section*{Acknowledgments}
\section*{References}
\begingroup
\renewcommand{\chapter}[2]{}
\bibliography{newbib.bib}
\endgroup
\bibliographystyle{plain}
